---
title: "04 Data Visualization"
author: "Vignesh Venkatachalam"
date: "2021-05"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE, align = "center", out.width = "110%")
```

# Data Visualization

### Load libraries

```{r load-libraries}
library(tidyverse)
library(ggthemes)
library(ggrepel)
library(lubridate)
library(maps)
```

## Challenge 1: Cumulative Cases

### Cumulative Cases Data Wrangling

I have decided not to add data for Europe and only add data by countries.

```{r data-wrangling-cum-cases}

covid_data_tbl <- read_csv("https://covid.ourworldindata.org/data/owid-covid-data.csv") %>% 
  
  select(location, date, total_cases) %>% 
  
  # Choosing data for 6 countries
  filter(location == "Germany" | location == "United Kingdom" | location == "France" |
         location == "Spain" | location == "United States" | location == "India") %>% 
  
  # Format cases data to decimal format
  mutate(cases_format = scales::dollar(total_cases, big.mark = ".",
                                       decimal.mark = ",",
                                       prefix = "",
                                       suffix = ""))
```

### Plot Cumulative Cases

The time course of cumulative cases is plotted below.

```{r plot-cum-cases}
covid_data_tbl %>% 
  
  ggplot(aes(x = date, y = total_cases, color = location)) +
  
  geom_line(size = 1) +
  
  # Used the ggrepel package to avoid data labels from overlapping
  geom_label_repel(aes(x = date, y = total_cases, label = cases_format, fill = location),
                   vjust = 0.7,
                   hjust = 1.2,
                   color = "white",
                   box.padding = 0.0,
                   min.segment.length = 0,
                   segment.color = "grey50",
                   data  = covid_data_tbl %>% slice(which.max(total_cases))) +
  
  # Removing unnecessary legend data
  guides(fill = FALSE) +
  
  expand_limits(y = 0) +
  
  scale_color_brewer(palette = "Dark2") +
  
  scale_y_continuous(labels = scales::dollar_format(scale = 1e-6,
                                                    prefix = "",
                                                    suffix = " M")) +
  
  scale_x_date(date_labels = "%B '%y",
               date_breaks = "1 month",
               expand = c(0, NA)) +
  
  # Labels
  labs(
    title = "COVID-19 confirmed cases worldwide",
    subtitle = "As of 08/05/2021",
    x = "",
    y = "Cumulative Cases",
    color = "Country"
  ) +
  
  # Themes
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1.0)
  )
```

## Challenge 2: Mortality Rate

### Mortality Rate Data Wrangling

```{r data-wrangling-mort-rate}
covid_data_tbl <- read_csv("https://covid.ourworldindata.org/data/owid-covid-data.csv") %>% 
  
  mutate(mortality_rate = total_deaths / population) %>% 
  
  # Getting the last mortality rate to plot
  group_by(location) %>% 
  summarise(latest_mort_rate = last(mortality_rate)) %>% 
  select(location, latest_mort_rate) %>% 
  
  # Taking care of inconsistencies in country names
  mutate(location = case_when(
    
    location == "United Kingdom" ~ "UK",
    location == "United States" ~ "USA",
    location == "Democratic Republic of Congo" ~ "Democratic Republic of the Congo",
    TRUE ~ location
    
  )) %>%
  
  distinct()

# Map data from the map package
world <- map_data("world")

covid_map <- merge(x = world, y = covid_data_tbl, by.x = "region", by.y = "location") %>% 
  select(region, long, lat, latest_mort_rate)

```

### Heat Map of Mortality Rate

The heat map of the mortality rates for the world is plotted below.

```{r heat-map-mort-rate}

covid_map %>% 
  ggplot() +
  
  geom_map(aes(x = long, y = lat, map_id = region, fill = latest_mort_rate), map = world) +
  
  # Add border lines for the countries
  borders("world", colour = "grey70") +
  
  # Colour gradient for heat map
  scale_fill_continuous(labels = scales::percent_format(accuracy = 0.001), 
                                                        low = "firebrick1", 
                                                        high = "darkred") +
  
  # Labels
  labs(
    title = "Confirmed COVID-19 deaths relative to the size of the population",
    subtitle = "Around 3 Million confirmed COVID-19 deaths worldwide",
    caption = "Date: 08/05/2021",
    fill = "Mortality Rate"
  ) +
  
  # Themes
  theme_minimal() +
  
  theme(
    axis.line = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    title = element_text(color = "black"),
    legend.position = "right"
  )
```